### 2017-01-08 project_1.Rmd
initial markdown commit

---
title: "project 1"
author: "Zijiang Yang, Dylan Sun"
date: "2017-01-08"
output: html_document
---

# Hello
```{r}

```