### 2017-01-24 final_presentation.Rmd
final presentation done

---
title: "Interim Presentation"
author: "Dylan Sun & Zijiang Yang"
date: '2017-01-17'
output:
  slidy_presentation: 
    font_adjustment: +2
---

```{r, message=FALSE, warning=FALSE, echo = FALSE, results = "hide"}
library(readxl)
library(knitr)
library(data.table)
library(corrplot)
library(broom)
library(stargazer)
library(ggplot2)
library(gridExtra)
library(ROCR)
prostate <- read_excel("~/Downloads/Prostate SBRT Sexual Function Data.xlsx", skip = 7)
prostate <- data.table(prostate)
prostate <- prostate[complete.cases(prostate)] # remove last row which is empty
# Code ADT as numeric; Y = 1, N = 0
prostate[ADT == "Y", ADT := 1]
prostate[ADT == "N", ADT := 0]
prostate[, ADT := as.numeric(ADT)]
prostate[, centered_age := Age - mean(Age)]
prostate[`Gleason Score` <= 6, `Gleason Score` := 0]
prostate[`Gleason Score` >= 8, `Gleason Score` := 2]
prostate[`Gleason Score` == 7, `Gleason Score` := 1]
prostate[`T-Stage Group` == 2, `T-Stage Group` := 1]
```

## Objectives
- Determine predictors of erectile function preservation after SBRT
- Determine the erectile function rates following SBRT
- Determine who is likely to have erectile function

## Potential Predictors
Negatively Associated:
- Age, Gleason Score, T-stage, PSA, ADT, BMI

Positively Associated:
- HRQOL, Erectile Function at Baseline

We expect HRQOL and Erectile Function at Baseline to be highly correlated, because erectile function is included in the questionnaire.

## How many people lose function?
- At baseline:
158 with function, 167 without (0.49)

- At two years:
102 with function, 223 without (0.31)

- A seemingly large number of people lose function in just two years

```{r echo=FALSE}
nf_nf <- dim(prostate[`Erectile Function at Baseline` == 0 & `Erectile Function at 2 years` == 0])[1]
nf_hf <- dim(prostate[`Erectile Function at Baseline` == 0 & `Erectile Function at 2 years` == 1])[1]
hf_nf <- dim(prostate[`Erectile Function at Baseline` == 1 & `Erectile Function at 2 years` == 0])[1]
hf_hf <- dim(prostate[`Erectile Function at Baseline` == 1 & `Erectile Function at 2 years` == 1])[1]
counts <- data.table(never_functional = nf_nf, gain_function = nf_hf, loss_function = hf_nf, retain_function = hf_hf)
kable(counts)
```

## Data processing
- Age was centered
- Gleason score was collapsed into "low" (GS <= 6), "medium" (GS == 7), and "high" (GS >= 8) 
- T-Stage group was collapsed into two categories, 0 and 1; the lone patient with T-Stage group >= 2 went into the 1 group.

## Model 
- Outcome variable: "Erectile Function at 2 Years" 
- Predictors: Age, Gleason Score, PSA, T-Stage Group, ADT, BMI, Erectile Function at Baseline
- Logistic regression using the glm method in R

## Results
```{r echo=FALSE, results = "hide"}
prostate_logreg <- glm(`Erectile Function at 2 years` ~ centered_age + `Gleason Score` + `T-Stage Group` + PSA + HRQOL + ADT + BMI + `Erectile Function at Baseline`, data = prostate, family = binomial(link = "logit"))

prostate_logreg_results <- data.table(tidy(prostate_logreg))
prostate_logreg_results[, lower := exp(estimate - 1.96*std.error)]
prostate_logreg_results[, OR := exp(estimate)]
prostate_logreg_results[, upper := exp(estimate + 1.96*std.error)]

prostate[, predicted_prob := predict(prostate_logreg, type = "response")]
```
```{r echo = FALSE, message=FALSE, warning=FALSE, results="asis"}
stargazer(prostate_logreg, header = F, ci=T, single.row = T, coef = list(prostate_logreg_results$OR), ci.custom = list(prostate_logreg_results[, .(lower, upper)]), p = list(prostate_logreg_results$p.value), keep = 1:8, keep.stat = c("n"), title = "Logistic regression results", dep.var.labels   = "Odds ratios for one unit change (95\\% CI)", dep.var.caption  = "\\it{For Erectile function at baseline:}", notes.align = "c", type = "html")
```

## Results
```{r ROC Curve, echo = FALSE, fig.cap="ROC Curve"}
pred <- prediction(prostate$predicted_prob, prostate$`Erectile Function at 2 years`)
roc.perf <- performance(pred, measure = "tpr", x.measure = "fpr")
# I know, the following code is bizarre. Just go with it.
auc <- performance(pred, measure = "auc")
auc <- auc@y.values[[1]]

roc.data <- data.frame(fpr=unlist(roc.perf@x.values),
                       tpr=unlist(roc.perf@y.values),
                       model="GLM")
ggplot(roc.data, aes(x=fpr, ymin=0, ymax=tpr)) +
    geom_ribbon(alpha=0.2) +
    geom_line(aes(y=tpr)) + 
   labs(x = "false positive rate", y = "true positive rate") +
  geom_abline(slope = 1)
```

## Error rates
- AUC = 0.842
- Threshold = 0.426 found using Euclidean distance
- 38 percent predicted to have erectile function at two years; from 49 percent at baseline (0.225 error rate)
- Using 0.5 as the cutoff, 30 percent expected to retain (0.206 error rate)
- In dataset, 31 percent of patients had function at two years
- Calculated threshold is more optimistic and encourages treatment

## Conclusions
- Age, HRQOL, and erectile function are predictive of erectile function at two years
- HRQOL and erectile function at baseline are protective, while age is detrimental
- PSA is somewhat protective
- Younger men, with high HRQOL scores and high PSA levels, and have erectile function at baseline will likely retain their erectile function 
- HRQOL and age are the most relevant; PSA levels less so.